---
title: "HIV Data Analysis Report"
author: "Birra Bejiga"
date: "2025-11-02"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r frequency}
# Load flextable package
library(flextable)
 # 1️⃣ Read your data
hiv <- read.csv("hiv.csv")

# 2️⃣ Create a frequency table
HIV_status_table <- table(hiv$HIV.status)
hiv
flextable::flextable(hiv)
#3️⃣ Create a percentage table
HIV_Status_Percent <- 100 * HIV_status_table /sum(HIV_status_table)
HIV_status_table 
```

```{r barplot_HIV_status}
barplot(HIV_Status_Percent)
```


```{r barplot_HIV_status_colered}
HIV_Status_Table <- table(hiv$HIV.status)
HIV_Status_Percent <- 100 * HIV_Status_Table / sum(HIV_Status_Table)




bp <- barplot(HIV_Status_Percent,
main = "HIV/AIDS Prevalence",
xlab = "HIV Status",
ylab = "Prevalence (%)",
col = c("green", "red"),
ylim = c(0, 100),
border = "blue",
density = c(10, 40))
abline(h=0)
```

#Bar Plot for HIV Status by Area
```{r Bar Plot for HIV Status by Area}
head(hiv)
# Barplot with grouped bars
HIV_by_area <- table(hiv$HIV.status, hiv$area)
# View the table
print(HIV_by_area)
# Convert to percentages if needed
HIV_by_area_percent <- 100*HIV_by_area/sum(HIV_by_area)
# View the percentages
print(round(HIV_by_area_percent, 1))
```


```{r Bar Plot for HIV Status by Area_Visualization}

# Assuming HIV_by_area_percent is a table or matrix with percentages
barplot(HIV_by_area_percent,
        main = "HIV Status by Residence",   # Title in quotes
        xlab = "Residence",                 # X-axis label
        ylab = "Prevalence (%)",            # Y-axis label
        col = c("green", "red"),            # Colors for bars
        beside = TRUE,                      # Side-by-side bars
        ylim = c(0, 50),                    # Y-axis limits
        xlim = c(1, 8))                     # X-axis limits

# Add horizontal line at y=0
abline(h = 0)
legend("topright", c("No","Yes"), fill = c("green","red"), title= "HIV Status")
HIV_by_area_percent
text(1.5, 25.5, "23.8%")
text(2.5,20.0, "19%")
text(4.5,25.5,"26.1%")
text(5.5,30.5,"31%")

